Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2018-12-13 mprop/mprop.tex
The proof is complete but messy

\begin{document}
\title{Variational Inference for Inverse Reinforcement Learning with Gaussian
  Processes}
\author{Paulius Dilkas}
\date{4th December 2018}
\maketitle
\tableofcontents
\newpage

@@

Following previous work on GP IRL
\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, we use a maximum
entropy IRL model \cite{ziebart2008maximum}, under which we have that
\[ P(a | s) \propto \exp(Q_r(s, a)), \]
where
\begin{equation}
  Q_r(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s, a, s')V_r(s'),
\end{equation}
and $V_r(s)$ is a `soft' version of the Bellman backup operator, which can be
obtained by repeatedly applying the following equation until convergence:
\cite{DBLP:conf/nips/LevinePK11,supplementary_material}
\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
The likelihood of the data can then be written down as
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
\begin{equation} \label{pDr}
  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right).
\end{equation}
However, a reward function learned by maximising this likelihood is not
transferable to new situations
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
reward structure in a way that would allow reward predictions for previously
unseen states.

@@

The work plan can be summarised as follows:
\begin{enumerate}
\item Derive expressions for the derivatives of the ELBO w.r.t. the variational
  parameters. \emph{Deliverables}: full derivations that can be included as
  supplementary material of the final paper, including formal proofs for
  applications of the dominated convergence theorem. \emph{Date}: 15~December.
\item Implement the algorithm as a more specialised version of Algorithm~2 in
  the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14}. As most of the relevant
  IRL algorithms have been implemented in
  MATLAB\footnote{\url{http://graphics.stanford.edu/projects/gpirl/irl_toolkit.zip}}\footnote{\url{https://github.com/jinming99/DGP-IRL}},
  it makes sense to do the same in order to have accurate time-sensitive
  comparisons between the algorithms. \emph{Deliverables}: a working
  implementation that can be integrated into the IRL toolkit by Levine et al.
  \cite{DBLP:conf/nips/LevinePK11}, and the same algorithm written in pseudo
  code for the paper. \emph{Date}: 1~February.
\item Evaluate the algorithm by comparing it to multiple alternatives, as
  detailed in Section \ref{sec:evaluation}. \emph{Deliverables}: multiple plots
  in an SVG or PDF format, along with code that produces them. \emph{Date}:
  1~April.
\item Write/finalise the paper. \emph{Deliverable}: a 14-page paper.
  \emph{Date}: 12~April.
\end{enumerate}

>>>>

\begin{document}
\title{Variational Inference for Inverse Reinforcement Learning with Gaussian
  Processes}
\author{Paulius Dilkas}
\date{12th December 2018}
\maketitle
\tableofcontents
\newpage

@@

Following previous work on GP IRL
\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, we use a maximum
entropy IRL model \cite{ziebart2008maximum}, under which we have that
\[ P(a | s) \propto \exp(Q_r(s, a)), \]
where
\begin{equation}
  Q_r(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s, a, s')V_r(s'),
\end{equation}
and $V_r : \mathcal{S} \to \mathbb{R}$ is an MDP \emph{value function}, which
can be obtained by repeatedly applying the 'soft' version of the Bellman backup
operator until convergence:
\cite{DBLP:conf/nips/LevinePK11,supplementary_material}
\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
The likelihood of the data can then be expressed as
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
\begin{equation} \label{pDr}
  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right).
\end{equation}
However, a reward function learned by maximising this likelihood is not
transferable to new situations
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
reward structure in a way that would allow reward predictions for previously
unseen states.

@@

The work plan can be summarised as follows:
\begin{enumerate}
\item Derive expressions for the derivatives of the ELBO w.r.t. the variational
  parameters. \emph{Deliverables}: full derivations that can be included as
  supplementary material of the final paper, including formal proofs for
  applications of the dominated convergence theorem. \emph{Date}: 15~January.
\item Implement the algorithm as a more specialised version of Algorithm~2 in
  the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14}. As most of the relevant
  IRL algorithms have been implemented in
  MATLAB\footnote{\url{http://graphics.stanford.edu/projects/gpirl/irl_toolkit.zip}}\footnote{\url{https://github.com/jinming99/DGP-IRL}},
  it makes sense to do the same in order to have accurate time-sensitive
  comparisons between the algorithms. \emph{Deliverables}: a working
  implementation that can be integrated into the IRL toolkit by Levine et al.
  \cite{DBLP:conf/nips/LevinePK11}, and the same algorithm written in pseudo
  code for the paper. \emph{Date}: 15~February.
\item Evaluate the algorithm by comparing it to multiple alternatives, as
  detailed in Section \ref{sec:evaluation}. \emph{Deliverables}: multiple plots
  in an SVG or PDF format, along with code that produces them. \emph{Date}:
  1~April.
\item Write/finalise the paper. \emph{Deliverable}: a 14-page paper.
  \emph{Date}: 12~April.
\end{enumerate}
